---
title: "p8105_hw3_cd3591"
author: "Anny"
date: "2025-10-10"
output: github_document
---

```{r}
# set up libraries and the dataset
library(tidyverse)
library(patchwork)
library(p8105.datasets)
data("instacart")
```

# Problem 1
```{r}
# basic structure of the dataset
str(instacart)
head(instacart) # illustrative examples of observations
```

**Short Description**: The instacart dataset contains `r nrow(instacart)` observations of `r ncol(instacart)` variables, representing individual grocery orders placed through the Instacart platform. Each observation corresponds to a single product ordered by a user. Key variables include: order_id, indicating unique identifier for each order;  product_id & product_name, identifying the product ordered; aisle_id / department_id, describing where the product is found; order_hour_of_day, showing the time of order placement; order_dow, indicating the day of week (0 = Sunday).

Q1
```{r}
# How many aisles are there, and which aisles are the most items ordered from?
aisle = instacart %>% count(aisle, sort = TRUE)
n_aisle = nrow(aisle)
head(aisle, 3)
```

There are `r n_aisle` aisles in total. The most frequently ordered aisles are fresh vegetables, fresh fruits, and packaged vegetables fruits.

Q2
```{r}
aisle %>%
  filter(n > 10000) %>%  # limiting this to aisles with more than 10000 items ordered
  mutate(aisle = fct_reorder(aisle, n)) %>%
  ggplot(aes(x = n, y = aisle)) +  # Make a plot that shows the number of items ordered in each aisle
  geom_col() +
  labs(title = "Number of Items Ordered per Aisle (>10,000 orders)",
       x = "Number of Items Ordered", y = "Aisle Name")   
       # since there are long labels like “packaged vegetables fruits”, fit much better along the y-axis than rotated on the x-axis.
```

Q3
```{r}
popular_items = instacart %>%  # Make a table
  filter(aisle == "baking ingredients" |
           aisle == "dog food care" |
           aisle == "packaged vegetables fruits") %>%  
  # in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”
  group_by(aisle, product_name) %>%
  summarise(times_ordered = n(), .groups = "drop_last") %>%  # Include the number of times each item is ordered in your table
  slice_max(times_ordered, n = 3)  # showing the three most popular items

popular_items
```

Q4
```{r}
mean_hour = instacart %>%  # Make a table
  filter(product_name == "Pink Lady Apples" | 
           product_name == "Coffee Ice Cream") %>%  # at which Pink Lady Apples and Coffee Ice Cream
  group_by(product_name, order_dow) %>%  # ordered on each day of the week
  summarise(mean_hour = mean(order_hour_of_day), .groups = "drop") %>%  # showing the mean hour of the day
  pivot_wider(names_from = order_dow, values_from = mean_hour) %>%  # produce a 2 x 8 table
  rename(Sun = "0", Mon = "1", Tue = "2", Wed = "3", Thu = "4", Fri = "5", Sat = "6")

mean_hour
```

# Problem 2
```{r}
# read the dataset
zip_info = read.csv("Zip Codes.csv", na = c("NA", ".", ""))%>% 
  janitor::clean_names() # clean names
zori = read.csv("Zip_zori_uc_sfrcondomfr_sm_month_NYC.csv", na = c("NA", ".", ""))%>% 
  janitor::clean_names() # clean names

# tidy zori
zori_new = zori %>% pivot_longer(
  x2015_01_31:x2024_08_31, 
  names_to = "date",
  values_to = "ZORI",
  names_prefix = "x"
  ) %>%  # tidy the dates
  mutate(date = ymd(str_replace_all(date, "_", "-")),
         zip_code  = as.integer(region_name)
         ) 

# collapse zip_info to one row per ZIP to avoid many-to-many
zip_info_one = zip_info %>%
  mutate(county = if_else(county == "New York", "Manhattan", county),
         zip_code = as.integer(zip_code)) %>%
  group_by(zip_code) %>%
  summarise(borough = first(county),
            neighborhood = str_c(sort(unique(neighborhood)), collapse = "; "),
            .groups = "drop") %>%
  dplyr::mutate(neighborhood = dplyr::na_if(neighborhood, ""))  # identify the empty is NA

#merge together
NY_zori = zori_new %>%
  left_join(zip_info_one,
            by = "zip_code") %>%
  dplyr::mutate(neighborhood = dplyr::na_if(neighborhood, "")) %>%  # identify the empty is NA
  arrange(zip_code, date) %>%
  relocate(zip_code, date, ZORI)
```

Q1
```{r}
ny_1501_2408 = NY_zori %>%
  filter(date >= ymd("2015-01-01"), date <= ymd("2024-08-31"))  # filter the date between January 2015 and August 2024

zip_counts = ny_1501_2408 %>%
  count(zip_code, name = "n_months")  # create the counts of zip table

n_116  = sum(zip_counts$n_months == 116)  # How many ZIP codes are observed 116 times?

n_lt10 = sum(zip_counts$n_months < 10)  #How many are observed fewer than 10 times?

# There are 116 months between January 2015 and August 2024. How many ZIP codes are observed 116 times? How many are observed fewer than 10 times? Why are some ZIP codes are observed rarely and others observed in each month?
```

There are `r n_116` ZIP codes observed 116 times and there are `r n_lt10` observed fewer than 10 times. Some ZIP codes don’t have many rental listings, like some small or mostly owner-occupied areas. When there isn’t enough data in a month, Zillow may leave it missing. By contrast, ZIP codes with consistent, high listing volume (dense residential areas) have enough data every month, so they’re observed in all 116 months.

Q2
```{r}
average_rental = NY_zori %>%
  mutate(year = year(date)) %>%
  group_by(borough, year) %>%  # showing the average rental price in each borough and year (not month)
  summarise(avg_rent = mean(ZORI, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = year, values_from = avg_rent) %>%
  arrange(borough) %>%
  knitr::kable()

average_rental
```

**Comments**: During 2015.01–2024.08, Manhattan is consistently the most expensive borough, followed by Brooklyn, Queens, and the Bronx, with Richmond slightly above the Bronx once data appear in 2020. Rents generally rose from 2015 to 2019, flattened around 2020–2021, and then rebounded sharply to new highs by 2024 in every borough. The surge after 2021 is the strongest in Manhattan, widening the gap between high- and lower-priced areas. Early years for Richmond are missing, which is why those cells show NaN.

Q3
```{r}
# Make a plot showing NYC Rental Prices within ZIP codes for all available years.
nyc_rental_plot = NY_zori %>%
  mutate(year = year(date)) %>%
  group_by(zip_code, borough, year) %>%
  summarise(rent = mean(ZORI, na.rm = TRUE), .groups = "drop") %>%
  ggplot(aes(x = year, y = rent, group = zip_code)) +
  geom_line(alpha = 0.25, na.rm = TRUE) +
  stat_summary(aes(group = borough), fun = median, geom = "line", linewidth = 1, na.rm = TRUE) +
  facet_wrap(~ borough, ncol = 3) +  # facilitate comparisons across boroughs
  labs(title = "NYC ZIP-Code Rental Price",
       x = "Year", y = "Rent (USD)")

nyc_rental_plot
```

**Comments**: From the graph, rents rose slowly from 2015 to 2019, flattened around 2020–2021, and then jumped sharply after 2021, especially in Manhattan. Each thin line indicates the rents of each zip code. Manhattan was always the most expensive and showed the widest spread across ZIP codes, indicating large within-borough differences. Brooklyn came next and also showed noticeable spread; Queens sat in the middle with more moderate levels and variation. The Bronx was consistently the lowest, with tighter clustering of ZIPs. Richmond appeared only from about 2020 onward and tracked just above the Bronx. The thick line (borough median) in each panel highlighted these patterns: rebounds pushed all borough medians to new highs by the end of 2021, and the gap between Manhattan and lower-priced boroughs widened.

Q4
```{r}
# Compute the average rental price within each ZIP code over each month in 2023
avg_2023 = NY_zori %>%
  filter(year(date) == 2023) %>%
  group_by(zip_code, borough) %>%
  summarise(avg_2023 = mean(ZORI, na.rm = TRUE), .groups = "drop")

# Make a reader-friendly plot showing the distribution of ZIP-code-level rental prices across boroughs
avg_2023_pplot = ggplot(avg_2023, aes(x = borough, y = avg_2023)) +  
  # facilitate the comparison of the distribution of average rental prices across boroughs
  geom_boxplot(outlier.alpha = 0.3, na.rm = TRUE) +
  labs(title = "Distribution of ZIP-Level Average Rent in 2023 by Borough",
       x = "Borough", y = "Average 2023 Rent (USD)")

avg_2023_pplot
```

**Comments**: In 2023, rents were the highest and most uneven in Manhattan, with several zip codes with very expensive rents pulling the upper tail up. Brooklyn followed with a lower median than Manhattan but still wide variation across ZIPs. Queens sat in the middle with moderate rents and a few higher-priced outliers. Richmond was a bit higher than the Bronx on median rent and showed the thinest box, suggesting uniform prices across its ZIP codes.

Q5
```{r}
# Combine the two previous plots into a single graphic
combine_plot = (nyc_rental_plot + avg_2023_pplot)

# export this to a results folder in your repository
dir.create("results", showWarnings = FALSE, recursive = TRUE)
ggsave("results/zillow_combined.png", width = 16, height = 8, dpi = 320)
```

# Problem 3
Q1
```{r}
# import the datasets
covar = read_csv("nhanes_covar.csv", na = c("NA", ".", "")) %>%
  janitor::clean_names()  # clean names
accel = read_csv("nhanes_accel.csv", na = c("NA", ".", "")) %>%
  janitor::clean_names()   # clean names

# tidy covar dataset
covar_fixed = covar %>%
  janitor::row_to_names(row_number = 4) %>%  # use row 4 as column names & remove the first 4 rows
  janitor::clean_names() %>%  # make nice: seqn, sex, age, bmi, education
  mutate(
    age = as.integer(age),
    bmi = as.numeric(bmi),
    seqn = as.character(seqn),
    sex = fct_recode(as.factor(sex),
                     "male"   = "1",
                     "female" = "2"),  # change 1 to male; 2 to female
    education = fct_recode(as.factor(education),
                           "less than high school" = "1", 
                           "high school equivalent" = "2", 
                           "more than high school" = "3")  
    #change 1 = less than high school; 2 = high school equivalent; 3 = more than high school
  )

# exclude participants less than 21 years of age, and those with missing demographic data
covar_clean = covar_fixed %>%
  filter(!is.na(seqn), !is.na(sex), !is.na(age), !is.na(education), age >= 21)

# tidy accel
accel = accel %>% 
  mutate(seqn = as.character(seqn))

# merge covar and accel
minute_cols = names(accel %>% select(starts_with("min")))

mims_data = covar_clean %>%
  inner_join(accel %>% select(seqn, all_of(minute_cols)), by = "seqn")

mims_data
```

Q2
```{r}
# Produce a reader-friendly table for the number of men and women in each education category
counts_tbl = covar_clean %>%
  count(education, sex, name = "n") %>%
  arrange(education, sex) %>% 
  pivot_wider(names_from = sex, values_from = n) %>%
  knitr::kable(caption = "Counts by education and sex (adults only)")

counts_tbl
  
# create a visualization of the age distributions for men and women in each education category
ggplot(covar_clean, aes(x = age, fill = sex)) +
  geom_histogram(position = "identity", alpha = 0.45, bins = 20) +
  facet_wrap(~ education, ncol = 3) +
  labs(title = "Age distributions by sex within education",
       x = "Age (years)", y = "Count", fill = "Sex") +
  theme_minimal()

```

**Comments**: From the table, men and women were about even in less than high school (28 vs 29) and in more than high school (56 vs 59), while high school equivalent was male-heavy (36 men vs 23 women). From the age plots, most participants had more than high school education, and they were mostly in their 20s–40s; men and women looked very similar in that group. The high school equivalent group was smaller and more spread out across ages, with no clear sex differences. Furthermore, the less than high school group was the smallest and tended to be older overall (fewer people under 40). Overall, the age patterns were driven more by education level than by sex, and the overlapping bars suggest only minor male–female differences within each category.

Q3
```{r}
# aggregate across minutes to create a total activity variable for each participant
totals = mims_data %>%
  mutate(total_activity = rowSums(across(all_of(minute_cols)), na.rm = TRUE))

# Plot these total activities (y-axis) against age (x-axis)
ggplot(totals, aes(x = age, y = total_activity, color = sex)) +  # compare men to women
  geom_point(alpha = 0.6) +
  geom_smooth(se = FALSE, method = "loess") +  # Include a trend line or a smooth to illustrate differences.
  facet_wrap(~ education, ncol = 3) +  # have separate panels for each education level
  labs(title = "Total daily activity (MIMS sum) vs Age",
       x = "Age", y = "Total MIMS", color = "Sex") +
  theme_minimal()
```

**Comments**: Across all three education groups, total daily activity generally went down as age went up. In more than high school and less than high school, the decline was steady from the 20s/30s into older ages. Moreover, the decline was more gentle in more than high school, compared to less than high school. In high school equivalent, there was a small mid-life bump (around the 30s–40s) before activity dropped in later years. Men and women looked very similar overall. In specific, their smooth lines sat close together, though women were a tad higher in activity in parts of the middle ages in the high school equivalent and more than high school panels. The main trend is the age-related decline, with only small sex differences.

Q4
```{r}
# tidy the mims_data for the preparing for the plot
mims_long = mims_data %>%
  select(seqn, sex, education, all_of(minute_cols)) %>%
  pivot_longer(all_of(minute_cols), names_to = "minute", values_to = "mims") %>%
  mutate(minute_num = readr::parse_number(minute)) %>%
  filter(dplyr::between(minute_num, 1, 1440))

prof = mims_long %>%
  group_by(education, sex, minute_num) %>%
  summarise(mims_mean = mean(mims, na.rm = TRUE), .groups = "drop")

# three-panel plot that shows the 24-hour activity time courses
ggplot(prof, aes(x = minute_num/60, y = mims_mean, color = sex)) +  # use color to indicate sex & change x into hours
  geom_line(alpha = 0.9) +
  geom_smooth(se = FALSE, method = "loess") +
  facet_wrap(~ education, ncol = 1) +  # for each education level
  scale_x_continuous(breaks = seq(0, 24, 4)) +
  labs(title = "Average 24-hour activity profiles by education",
       x = "Time of day (hour)", y = "Mean MIMSC:/Users/13013/OneDrive/桌面/BIST8105P data sciencec I/p8105_hw3_cd3591/nhanes_accel.csv", color = "Sex") +
  theme_minimal()
```

**Comments**: Across all three education groups, activity was the lowest overnight, then rose sharply in the morning, peaked around mid-day till early afternoon, and then gradually declined into the evening. The daily shape was very similar everywhere. The highest mid-day levels appeared in the less than high school group and more than high school, the high school equivalent group was close behind. Men and women followed nearly the same pattern; any sex differences were small. In specific, women tended to be slightly higher during the day in the high school equivalent and more than high school groups, while men were a touch higher very early in the morning in the less than high school group.