---
title: "p8105_hw3_cd3591"
author: "Anny"
date: "2025-10-10"
output: github_document
---

```{r}
# set up libraries and the dataset
library(tidyverse)
library(p8105.datasets)
data("instacart")
```

# Problem 1
```{r}
# basic structure of the dataset
str(instacart)
head(instacart) # illustrative examples of observations
```
*Short Description*: The instacart dataset contains `r nrow(instacart)` observations of `r ncol(instacart)` variables, representing individual grocery orders placed through the Instacart platform. Each observation corresponds to a single product ordered by a user. Key variables include: order_id, indicating unique identifier for each order;  product_id & product_name, identifying the product ordered; aisle_id / department_id, describing where the product is found; order_hour_of_day, showing the time of order placement; order_dow, indicating the day of week (0 = Sunday).

1)
```{r}
# How many aisles are there, and which aisles are the most items ordered from?
aisle = instacart %>% count(aisle, sort = TRUE)
n_aisle = nrow(aisle)
head(aisle, 3)
```
There are `n_aisle` aisles in total. The most frequently ordered aisles are fresh vegetables, fresh fruits, and packaged vegetables fruits.